---
title: 'Summary Stats on the New Data Dump: Mark II'
output: html_notebook
date: '2017-06-07'
---

Let's do some quick summary stats on the new data dump, after removing the duplicates I discovered - see `5_31_sum_stats.Rmd`.

```{r setup}
library(readr)
library(dplyr)
library(ggplot2)
library(ggTimeSeries)
library(lubridate)
library(geosphere)
library(viridis)
library(stringr)
library(lme4)

raw_locations <- read_rds('../data/raw/Michigan_DB_user_location_05_31_17.rds')
cleaned_locations <- read_rds('../data/processed/Michigan_DB_user_location_05_31_17_cleaned.rds')
std_locations <- read_rds('../data/processed/std_locations.rds')
twin_distances <- read_rds('../data/processed/twin_distances.rds')
id_mapping <- read_csv('../data/processed/id_mapping_long.csv', col_types = 'ccc')
user_data <- read_rds('../data/raw/Michigan_DB_users_05_31_17.rds')
cleaned_twin_locations <- inner_join(cleaned_locations, id_mapping, by = c("user_id" = "alternate_id")) %>% left_join(user_data, c("user_id" = "alternate_id"))
tweets <- read_rds('../data/raw/Michigan_DB_user_tweet_05_31_17.rds') %>% mutate(tweet_id = as.character(tweet_id))
checkin <- read_rds('../data/processed/remote_substance_use.rds') %>% inner_join(id_mapping, by = c('user_id' = 'alternate_id'))
twin_info <- read_rds('../data/raw/Robin_paper-entry_2-22-17.rds')
par_mon <- read_rds('../data/processed/remote_parents.rds')

comma <- function(x) format(x, digits = 2, big.mark = ",")
```


# Location

In the row data, we have `r comma(nrow(raw_locations))` points in the current data dump. After deduplicating, filtering out observations with missing fields,  sample_time values outside of the appropriate ranges, points with an accuracy greater than or equal to 500 meters, and points that don’t correspond to a twin, we have `r comma(nrow(cleaned_twin_locations))` on `r length(unique(cleaned_twin_locations))` individuals. Of those individuals, `r user_data %>% filter(app_type == 'ios') %>% nrow()` have an iPhone, `r user_data %>% filter(app_type == 'android') %>% nrow()` have an Android phone, and `r sum(is.na(user_data$app_type))` do not have a recorded device type.

Let's visualize the distribution of the number of points per twin.

```{r}
cleaned_twin_locations %>% group_by(user_id) %>% summarize(n = n()) %>% filter(n < 50000) %>% left_join(user_data, by = c("user_id" = "alternate_id")) %>% ggplot(aes(x = n, color = app_type)) + geom_density() + xlab("Number of locations recorded")
```

The mean density of location points for iOS vs. Android. We’ll need the length of time between first and last locations for each twin, converting from seconds to days.

```{r}
first_last_locs <- cleaned_twin_locations %>% group_by(user_id) %>% summarize(first_point = min(sample_time), last_point = max(sample_time)) %>% mutate(duration_days = as.numeric(last_point - first_point)/60/60/24)
cleaned_twin_locations %>% group_by(user_id) %>% summarize(n = n()) %>% left_join(user_data, by = c("user_id" = "alternate_id")) %>% left_join(first_last_locs) %>% mutate(loc_density = n / duration_days) %>% ggplot(aes(x = loc_density, color = app_type)) + geom_density() + xlab('Points Per Day') + coord_cartesian(xlim = c(0, 500))
```

Now let's look at the distribution of time between points.

```{r}
results <- tibble()
for (twin in unique(cleaned_twin_locations$user_id)) {
  results2 <- cleaned_twin_locations %>% filter(user_id == twin) %>% arrange(sample_time) %>% transmute(point_interval = lead(sample_time) - sample_time, app_type = app_type)
  results <- bind_rows(results, results2)
}
```

```{r}
results %>% mutate(point_interval = point_interval + 1) %>% ggplot(aes(x = point_interval, fill = app_type)) + geom_histogram() + scale_x_log10() + xlab('Interval in seconds')
```

Let's try looking at physical distance between points.

```{r}
results <- tibble()
for (twin in unique(cleaned_twin_locations$user_id)) {
  results2 <- cleaned_twin_locations %>% filter(user_id == twin) %>% arrange(sample_time) %>% transmute(point_interval = distHaversine(cbind(lead(.$longitude), lead(.$latitude)), cbind(.$longitude, .$latitude)), app_type = app_type)
  results <- bind_rows(results, results2)
}
```

```{r}
results %>% mutate(point_interval = point_interval + 1) %>% ggplot(aes(x = point_interval, fill = app_type)) + geom_histogram() + scale_x_log10() + xlab('Interval in meters')
```

Let's look at how many points we've received over time.

```{r}
cleaned_twin_locations %>% mutate(date = as.Date(sample_time)) %>% group_by(date) %>% summarize(n = n()) %>% ggplot_calendar_heatmap('date', 'n') + xlab(NULL) + ylab(NULL) + scale_fill_viridis() + facet_wrap(~Year, ncol = 1)
```

What device types have these points come from?

```{r}
cleaned_twin_locations %>% group_by(app_type) %>% summarize(n = n())
```


# Tweets

We haven't received tweet data since April 11. Bizarrely, the user ID is sometimes the hashed Colorado ID and sometimes the Michigan ID. Nevertheless, we have `r comma(nrow(tweets))` tweets from `r length(unique(tweets$alternate_id))` ID's. `r comma(sum(str_detect(tweets$text, 'RT')))` are retweets.

# Checking In Survey

The "Checking In" substance use survey has been completed `r comma(nrow(checkin))` times by `r length(unique(checkin$user_id))` twins. The distribution of number of completed surveys looks like this:

```{r}
checkin %>% group_by(user_id) %>% summarise(n = n()) %>% ggplot(aes(x = n)) + geom_histogram(binwidth = 5) + xlab('Number of completed checking in surveys')
```

```{r}
checkin %>% group_by(user_id) %>% summarise(n = n(), first = min(date_completed)) %>% ggplot(aes(x = first, y = n)) + geom_point() + geom_smooth(method = "lm") + xlab("Date of first checking in survey completed") + ylab("Number of completed checking in surveys")
```

In what proportion of weeks do twins use substances?

```{r}
checkin %>% group_by(user_id) %>% summarize(use_prop = sum(any_substance_use) / n()) %>% ggplot(aes(x = use_prop)) + geom_histogram(binwidth = 0.05) + xlab("Proportion of weeks in which any substance was used")
```


# Parental Monitoring Survey

The parental monitoring survey has been completed `r comma(nrow(par_mon))` times by `r length(unique(par_mon$user_id))` twins. The distribution of completed surveys looks like this.

```{r}
par_mon %>% group_by(user_id) %>% summarise(n = n()) %>% ggplot(aes(x = n)) + geom_histogram(binwidth = 1) + xlab('Number of completed parental monitoring surveys')
```

```{r}
par_mon %>% group_by(user_id) %>% summarise(n = n(), first = min(date_completed)) %>% ggplot(aes(x = first, y = n)) + geom_point(alpha = 0.5) + geom_smooth(method = "lm") + xlab("Date of first parental monitoring survey completed") + ylab("Number of completed parental monitoring surveys")
```


Let's look at the trajectories of parental monitoring score over the study for each twin.

```{r}
par_mon %>% ggplot(aes(x = date_completed, y = max_monitor_score, group = user_id)) + geom_line(alpha = 0.2) + xlab("Date of Survey Completion") + ylab("Parental Monitoring Score")
```
We can see that there is a tendency to persist at the current level or to shift to a closer level over a further level but that there is significant variation.

```{r}
data <- left_join(checkin, twin_info, by = c("SVID" = "ID1"))
data2 <- tibble()
for (twin in unique(data$user_id)) {
  t_data <- data[data$user_id == twin,]
  d   <- function(x,y) abs(x-y) # define the distance function
  idx <- sapply(t_data$date_completed, function(x) which.min( d(x,par_mon$date_completed) )) # find matches
  t_data <- cbind(t_data,par_mon[unlist(idx),-1,drop=FALSE])
  data2 <- bind_rows(data2, t_data)
}
```

```{r}
glmer(any_substance_use ~ Test_AgeYears + Sex1 + bestzygos.x + max_monitor_score + (1 | family/user_id), data = data2)
```


