---
title: "COTWINS Grant Data"
output: html_notebook
---

```{r, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)

paper_data <- read_rds("../data/processed/Robin_paper-entry_2-22-17_cleaned.rds")
phenx_diagnoses <- read_rds("../data/processed/PhenX_diagnoses.rds")
remote_sub_use <- read_rds("../data/processed/remote_substance_use.rds")
in_person_par <- read_rds("../data/processed/in_person_parents.rds")
remote_par <- read_rds("../data/processed/remote_parents.rds")
id_mapping_long <- read_csv("../data/processed/id_mapping_long.csv", col_types = "ccc")
id_mapping <- read_csv("../data/processed/id_mapping.csv", col_types = "ccccc")
locs <- read_rds("../data/processed/Michigan_DB_user_location_05_31_17_cleaned.rds")

comma <- function(x) format(x, digits = 3, big.mark = ",")
```

# Descriptives

There are `r nrow(id_mapping_long)` twins in the study. `r length(unique(remote_sub_use$user_id))` have completed at least one checking in survey. `r length(unique(locs$user_id[locs$user_id %in% id_mapping_long$alternate_id]))` have at least one GPS point after data cleaning. 

## Followup Rates

```{r}
checkin_rate <-
  remote_sub_use %>% group_by(user_id) %>% summarize(n = n()) %>% left_join(id_mapping_long, by = c("user_id" = "alternate_id")) %>% left_join(paper_data, by = c("SVID" = "ID1")) %>% mutate(time_in_study = as.duration(ymd("2017-05-31") - Test_Date)) %>% mutate(time_in_study = if_else(time_in_study <= dyears(1), time_in_study, dyears(1))) %>% mutate(
  weeks_in_study = as.numeric(time_in_study) / 604800,
  checkin_per_week = n / (as.numeric(time_in_study) / 604800)
  ) %>% select(user_id,
  n,
  Test_Date,
  time_in_study,
  checkin_per_week,
  weeks_in_study)
  
  ggplot(checkin_rate, aes(x = weeks_in_study, y = checkin_per_week)) + geom_point(alpha = 0.5)
```

`r comma(sum(checkin_rate$checkin_per_week > 0.5) / nrow(checkin_rate) * 100)`% of twins have a Checking In completion rate of greater than 50%.

## Sample Composition

```{r}
paper_data %>% mutate(Sex1 = as.numeric(Sex1)) %>% mutate(Sex1 = recode(Sex1, `1` = "Female", `2` = "Male")) %>% group_by(Zygosity = bestzygos, Sex = Sex1) %>% summarize(N = n(), `Mean Age at Assessment` = comma(mean(as.numeric(Test_Age / 365))))
```

## Birth Years

```{r}
paper_data %>% group_by(Birth_Year) %>% summarize(N = n())
```

# Validation of Remote Measures

```{r}
all_par_mon <- left_join(in_person_par, id_mapping_long, by = "SVID") %>% left_join(distinct(remote_par, user_id, .keep_all = T), by = c("alternate_id" = "user_id"))
cor(all_par_mon$max_monitor_score.x, all_par_mon$max_monitor_score.y, use = "complete.obs")
```
```{r}
lagged_remote <- remote_par %>% group_by(user_id) %>% summarize(first_mon = first(max_monitor_score), second_mon = nth(max_monitor_score, 2))
cor(lagged_remote$first_mon, lagged_remote$second_mon, use = "complete.obs")
```

# Twin Discordance

```{r}
tp_phenx <- left_join(id_mapping, phenx_diagnoses, by = c("T1" = "SVID")) %>% left_join(phenx_diagnoses, by = c("T2" = "SVID"), suffix = c(".T1", ".T2"))
```

Out of `r nrow(id_mapping_long)` twin pairs, `r sum(tp_phenx$alc_lifetime_use.T1 != tp_phenx$alc_lifetime_use.T2)` are discordant for lifetime alcohol use. `r sum(phenx_diagnoses$alc_lifetime_use)` twins have ever used alcohol. `r sum(tp_phenx$life_alc_depend.T1 != tp_phenx$life_alc_depend.T2)` twin pairs are discordant for a lifetime diagnosis of alcohol dependence. `r sum(phenx_diagnoses$life_alc_depend)` twins have a lifetime diagnosis of alcohol dependence.
