---
title: "COTWINS Grant Data"
output: html_notebook
---

```{r, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)

paper_data <- read_rds("../data/processed/Robin_paper-entry_2-22-17_cleaned.rds")
phenx_diagnoses <- read_rds("../data/processed/PhenX_diagnoses.rds")
remote_sub_use <- read_rds("../data/processed/remote_substance_use.rds")
in_person_par <- read_rds("../data/processed/in_person_parents.rds")
remote_par <- read_rds("../data/processed/remote_parents.rds")
id_mapping_long <- read_csv("../data/processed/id_mapping_long.csv", col_types = "ccc")
id_mapping <- read_csv("../data/processed/id_mapping.csv", col_types = "ccccc")
locs <- read_rds("../data/processed/Michigan_DB_user_location_05_31_17_cleaned.rds")

comma <- function(x) format(x, digits = 3, big.mark = ",")

theme_set(theme_gray(base_size = 13))
```

# Descriptives

There are `r nrow(id_mapping_long)` twins in the study. `r length(unique(remote_sub_use$user_id))` have completed at least one checking in survey. `r length(unique(locs$user_id[locs$user_id %in% id_mapping_long$alternate_id]))` have at least one GPS point after data cleaning. 

### Followup Rates

```{r}
checkin_rate <-
  remote_sub_use %>%
  group_by(user_id) %>%
  summarize(n = n()) %>%
  left_join(id_mapping_long, by = c("user_id" = "alternate_id")) %>%
  left_join(paper_data, by = c("SVID" = "ID1")) %>%
  mutate(time_in_study = as.duration(ymd("2017-05-31") - Test_Date)) %>%
  mutate(time_in_study = if_else(time_in_study <= dyears(1), time_in_study, dyears(1))) %>%
  mutate(
  weeks_in_study = as.numeric(time_in_study) / 604800,
  checkin_per_week = n / (as.numeric(time_in_study) / 604800)
  ) %>%
  select(
    user_id,
    n,
    Test_Date,
    time_in_study,
    checkin_per_week,
    weeks_in_study)
  
  ggplot(checkin_rate, aes(x = weeks_in_study, y = checkin_per_week)) + geom_point(alpha = 0.5)
```

`r comma(sum(checkin_rate$checkin_per_week > 0.5) / nrow(checkin_rate) * 100)`% of twins have a Checking In completion rate of greater than 50%.

### Sample Composition

```{r}
paper_data %>%
  mutate(Sex1 = as.numeric(Sex1)) %>%
  mutate(Sex1 = recode(Sex1, `1` = "Female", `2` = "Male")) %>%
  group_by(Zygosity = bestzygos, Sex = Sex1) %>%
  summarize(
    N = n(), 
    `Mean Age at Assessment` = comma(mean(as.numeric(Test_Age / 365)))
    )
```

### Birth Years

```{r}
paper_data %>%
  group_by(Birth_Year) %>%
  summarize(N = n())
```

# Validation of Remote Measures

The correlation between the in-person parental monitoring score and the first remote parental monitoring score for an individual.
```{r}
all_par_mon <-
  left_join(in_person_par, id_mapping_long, by = "SVID") %>%
  left_join(
    distinct(remote_par, user_id, .keep_all = T),
    by = c("alternate_id" = "user_id")
    )

cor(all_par_mon$max_monitor_score.x, all_par_mon$max_monitor_score.y, use = "complete.obs")
```

The correlation between the first and second remote parental monitoring scores for an individual.
```{r}
lagged_remote <-
  remote_par %>%
  group_by(user_id) %>%
  summarize(
    first_mon = first(max_monitor_score),
    second_mon = nth(max_monitor_score, 2)
    )

cor(lagged_remote$first_mon, lagged_remote$second_mon, use = "complete.obs")
```

The correlations are very similar, suggesting that any change due to the nature of the assessment is small compared to the week to week changes in the score.

# Twin Discordance

```{r}
tp_phenx <- 
  left_join(id_mapping, phenx_diagnoses, by = c("T1" = "SVID")) %>%
  left_join(phenx_diagnoses, by = c("T2" = "SVID"), suffix = c(".T1", ".T2"))
```

Out of `r nrow(id_mapping_long)` twin pairs, `r sum(tp_phenx$alc_lifetime_use.T1 != tp_phenx$alc_lifetime_use.T2)` are discordant for lifetime alcohol use. `r sum(phenx_diagnoses$alc_lifetime_use)` twins have ever used alcohol. `r sum(tp_phenx$life_alc_depend.T1 != tp_phenx$life_alc_depend.T2)` twin pairs are discordant for a lifetime diagnosis of alcohol dependence. `r sum(phenx_diagnoses$life_alc_depend)` twins have a lifetime diagnosis of alcohol dependence.

# Substance Use Trajectories

Get the subject's age in years at survey completion for all of the remote substance use responses.

```{r}
remote_sub_use_age <-
  remote_sub_use %>%
  left_join(id_mapping_long, by = c("user_id" = "alternate_id")) %>%
  left_join(paper_data, by = c("SVID" = "ID1")) %>%
  mutate(age_at_completion = as.numeric(as_date(date_completed) - Birth_Date) / 365)
```

If the subject claimed no substance use or no use of the substance in question in a given response, set the quantity consumed or frequency of use to zero.

### Alcohol

```{r}
remote_sub_use_age$alc_quantity_drinks_per_day[remote_sub_use_age$any_substance_use == FALSE] = 0
remote_sub_use_age$alc_quantity_drinks_per_day[remote_sub_use_age$alc_use == FALSE] = 0
ggplot(remote_sub_use_age, aes(x = age_at_completion, y = alc_quantity_drinks_per_day)) + geom_smooth() + xlab("Age in Years") + ylab("Drinks per Day") + xlim(14.5, 19)
```

### Marijuana

```{r}
remote_sub_use_age$mar_freq_times_per_day[remote_sub_use_age$any_substance_use == FALSE] = 0
remote_sub_use_age$mar_freq_times_per_day[remote_sub_use_age$mar_use == FALSE] = 0
ggplot(remote_sub_use_age, aes(x = age_at_completion, y = mar_freq_times_per_day)) + geom_smooth() + xlab("Age in Years") + ylab("Marijuana Use, Times per Day") + xlim(14.5, 19)
```

### Cigarettes

```{r}
remote_sub_use_age$cig_quantity_per_day[remote_sub_use_age$any_substance_use == FALSE] = 0
remote_sub_use_age$cig_quantity_per_day[remote_sub_use_age$cig_use == FALSE] = 0
ggplot(remote_sub_use_age, aes(x = age_at_completion, y = cig_quantity_per_day)) + geom_smooth() + xlab("Age in Years") + ylab("Cigarettes per Day") + xlim(14.5, 19)
```

### E-Cigarettes

```{r}
remote_sub_use_age$ecig_quantity_puffs[remote_sub_use_age$any_substance_use == FALSE] = 0
remote_sub_use_age$ecig_quantity_puffs[remote_sub_use_age$ecig_use == FALSE] = 0
ggplot(remote_sub_use_age, aes(x = age_at_completion, y = ecig_quantity_puffs)) + geom_smooth() + xlab("Age in Years") + ylab("E-Cigarette Puffs per Use") + xlim(14.5, 19)
```

```{r}
remote_sub_use_age$ecig_freq_times_per_day[remote_sub_use_age$any_substance_use == FALSE] = 0
remote_sub_use_age$ecig_freq_times_per_day[remote_sub_use_age$ecig_use == FALSE] = 0
ggplot(remote_sub_use_age, aes(x = age_at_completion, y = ecig_freq_times_per_day)) + geom_smooth() + xlab("Age in Years") + ylab("E-Cigarette Uses per Day") + xlim(14.5, 19)
```

