---
title: "Preliminary Analyses of the Colorado PhenX Substance Use Data"
output: html_notebook
date: 2017-01-31
---

```{r setup}
library(ggplot2)
library(readr)
library(dplyr)
library(broom)
library(forcats)

# The 'labelled' class columns need to be converted to standard data types
phenx <- read_rds('../data/raw/Robin_PhenX_12-6-16.rds')
phenx$id <- as.character(phenx$id)
ever_drinkers <- phenx %>% filter(PHXQ0001 == 1)
drinkers <- phenx %>% filter(PHXQ0002 == 1)

twin_info <- read_rds('../data/raw/Robin_paper-entry_12-6-16.rds')

twin_info$ID1 <- as.character(twin_info$ID1)
twin_info$Test_AgeYears <- as.numeric(twin_info$Test_AgeYears)

# Give sex more useful levels
twin_info$Sex1 <- fct_recode(as.factor(twin_info$Sex1),
                             'Male' = '1',
                             'Female' = '2')

# Join the data frames on SVID
joined <- left_join(phenx, twin_info, by = c('id' = 'ID1'))

# A function to format floats
comma <- function(x) format(x, digits = 2, nsmall = 2, big.mark = ",")
```


In the initial in-person COTWINS assessments, twins were asked standardized questions about substance use, derived from the [PhenX Toolkit](https://www.phenxtoolkit.org/index.php?pageLink=browse.conceptualgroups&id=3354&breadcrumbs=3354). The goal of this notebook is to do some preliminary descriptive analyses and QC on these data.

# Alcohol

Out of `r nrow(phenx)` twins, `r nrow(ever_drinkers)` admit to having at least one drink of alcohol. `r nrow(drinkers)` admit to having had more than five drinks. The mean age of drinking initiation was `r comma(mean(ever_drinkers$PHXQ0003))`, with this distribution, which shows some obvious outliers:

```{r Age of drinking initiation, warning=FALSE}
joined %>%
  ggplot() + 
  geom_histogram(aes(x = PHXQ0003, fill = Sex1), binwidth = 1) +
  labs(fill = 'Sex', x = 'Age of Drinking Intiation')
```

And the age at which participants began drinking once a month or more:

```{r Age of regular drinking}
joined %>%
  filter(PHXQ0004 != 0) %>%
  ggplot() + 
  geom_histogram(aes(x = PHXQ0004, fill = Sex1), binwidth = 1) +
  labs(fill = 'Sex', x = 'Age of Regular Drinking')
```
We observe that the mean age of ever drinkers is 16.08 and that of never drinkers is 15.26. Summaries of other questions follow.

During the past 30 days, on how many days did you have one or more drinks?

```{r Days in past month, warning=FALSE}
joined %>%
  ggplot() + 
  geom_histogram(aes(x = PHXQ0005, fill = Sex1), binwidth = 1) + 
  labs(fill = 'Sex', x = 'Days of Drinking')
```

How does the number of days of drinking vary with age?

```{r Days in past month x age, warning=FALSE}
joined %>%
  ggplot(aes(x = Test_AgeYears, y = PHXQ0005, color = Sex1)) + 
  geom_jitter() +
  labs(color = 'Sex', x = 'Age in Years', y = 'Days of Drinking')
```

If we fit a linear model, we can see that age in years is not significantly correlated with days of drinking.

```{r Days of drinking model}
glance(lm(PHXQ0005 ~ Test_AgeYears, data = joined))
```


During the past 30 days, when you drank, how many drinks did you have?

```{r Drinks per session, warning=FALSE}
joined %>%
  ggplot() + geom_histogram(aes(x = PHXQ0009, fill = Sex1), binwidth = 1) +
  labs(fill = 'Sex', x = 'Drinks per Session') +
  scale_x_continuous(breaks = seq(1, 12))
```

How does the number of drinks per session vary with age?

```{r Drinks per session x age, warning=FALSE}
joined %>%
  ggplot(aes(x = Test_AgeYears, y = PHXQ0009, color = Sex1)) + 
  geom_jitter() +
  labs(color = 'Sex', x = 'Age in Years', y = 'Drinks per Session')
```

If we fit a linear model, we can see that age in years is not significantly correlated with number of drinks per session.

```{r Drinks per session model}
glance(lm(PHXQ0009 ~ Test_AgeYears, data = joined))
```

